---
output: github_document
always_allow_html: true
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

# ilostat-wage

<!-- badges: start -->

[![License: CC BY
4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

This repository contains wage statistics data from the International Labour Organization (ILO), specifically focusing on average monthly earnings of employees disaggregated by sex and urban/rural areas. The data has been processed following the openwashdata publishing guidelines to provide a tidy dataset suitable for analysis.

## Installation

You can download the dataset as a CSV file from the table below.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)

extdata_path <- "https://github.com/openwashdata-conf/ilostat-wage/raw/main/"

data.frame(
  dataset = "ilostat_wage",
  CSV = paste0("[Download CSV](", extdata_path, "ilostat_wage.csv)")
) |> 
  knitr::kable()
```

## Data

The package provides access to ILO wage statistics data processed from the ILO Statistics Database.

```{r}
library(readr)
library(dplyr)
library(gt)

# Read the dataset
ilostat_wage <- read_csv("ilostat_wage.csv")
```

### ilostat_wage

The dataset `ilostat_wage` contains wage statistics from various countries processed from ILO Statistics data.
It has `r nrow(ilostat_wage)` observations and `r ncol(ilostat_wage)` variables.

```{r}
ilostat_wage |> 
  head(3) |> 
  gt::gt() |>
  gt::as_raw_html()
```

For an overview of the variable names, see the following table.

```{r echo=FALSE, message=FALSE, warning=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::select(variable_name:description) |> 
  knitr::kable() |> 
  kableExtra::kable_styling("striped") |> 
  kableExtra::scroll_box(height = "200px")
```

## Dataset Description

Dataset characteristics:

- **Source**: ILO Statistics Database
- **Indicator**: Average monthly earnings of employees
- **Currency**: 2021 PPP $ (Purchasing Power Parity adjusted US Dollars)
- **Disaggregation**: By sex (Total, Male, Female) and area type (National, Rural, Urban)
- **Time Coverage**: Various years depending on country data availability
- **Filter Date**: 2025-06-01

## Example

```{r}
library(ggplot2)
library(tidyverse)

# Distribution of earnings by sex
ilostat_wage |>
  filter(sex != "Total", !is.na(earnings_ppp)) |>
  ggplot(aes(x = earnings_ppp, fill = sex)) +
  geom_histogram(alpha = 0.7, position = "identity", bins = 30) +
  scale_x_log10() +
  labs(title = "Distribution of Monthly Earnings by Sex",
       x = "Monthly Earnings (2021 PPP $, log scale)",
       y = "Frequency",
       fill = "Sex") +
  theme_minimal()
```

### Which countries have the highest wage gaps?

```{r}
# Calculate wage gaps by country
wage_gaps <- ilostat_wage |>
  filter(sex %in% c("Male", "Female"), 
         area_type == "National",
         !is.na(earnings_ppp)) |>
  group_by(country, year) |>
  filter(n() == 2) |>  # Only countries with both male and female data
  summarise(
    male_earnings = earnings_ppp[sex == "Male"],
    female_earnings = earnings_ppp[sex == "Female"],
    wage_gap = (male_earnings - female_earnings) / male_earnings * 100,
    .groups = "drop"
  ) |>
  arrange(desc(wage_gap))

wage_gaps |>
  head(10) |>
  gt::gt() |>
  gt::fmt_number(columns = c(male_earnings, female_earnings), decimals = 0) |>
  gt::fmt_number(columns = wage_gap, decimals = 1, suffix = "%") |>
  gt::as_raw_html()
```

### Urban vs Rural wage differences

```{r}
# Compare urban vs rural wages
ilostat_wage |>
  filter(area_type %in% c("Urban", "Rural"),
         sex == "Total",
         !is.na(earnings_ppp)) |>
  ggplot(aes(x = area_type, y = earnings_ppp, fill = area_type)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(title = "Urban vs Rural Monthly Earnings Distribution",
       x = "Area Type",
       y = "Monthly Earnings (2021 PPP $, log scale)",
       fill = "Area Type") +
  theme_minimal()
```

## License

Data are available as
[CC-BY](https://github.com/openwashdata-conf/ilostat-wage/blob/main/LICENSE).

## Citation

Please cite this dataset using:

International Labour Organization. ILOSTAT database. Available from https://ilostat.ilo.org/data/.

For this processed dataset, please also cite:
[Author]. (2025). ILO Wage Statistics Dataset. Retrieved from https://github.com/openwashdata-conf/ilostat-wage